Simplifying Lotka--Volterra interaction matrices by species grouping: optimal groupings are unpredictable

### 2022-05-22 article.rnw
Move the word "however" from the middle to the beginning of a sentence.
Remove some commented out sentences.
Add a concluding "what to do going forward" sentence to the abstract.

Dynamic models of ecological communities require many parameters, making them expensive to fit to experimental or observational data.
To reduce the number of parameters, species are often divided into groups \emph{a priori}, using phylogenetic or functional groups, and species within a group are assumed to behave identically.
Here, we assess the validity of such grouping by deriving the optimal groupings \emph{a posteriori}, from the parsimony of a Lotka--Volterra model fit to experimental data from six species of \emph{Drosophila} competing in laboratory conditions.
We find evidence that grouping can improve the parsimony of species interaction models, but we also observe that there is no single optimal grouping.
Further, those groupings that did prove beneficial could not be easily predicted from available species attributes \emph{a priori}, calling into question the prevailing method of \emph{a priori} species grouping.
Lastly, we find that species did not group equivalently in their competitive effect and in their response to competition, invalidating a key assumption often made implicitly when defining species groupings.
% DBS: add some sort of concluding sentence a la where to from here

@@

The Lotka--Volterra competition model in Eq.~(\ref{LVmodel}) is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, we instead fit the models with the Gauss--Newton method for non-linear least squares and assumed a normal error distribution of the response variable.
%The models were fitted to the data using maximum likelihood regression techniques.
%Fitting it using precisely a Lotka--Volterra model prevented the use of a link function, rendering most generalised regression techniques infeasible.
To prevent heteroskedasticity and improve normality, we transformed total population growth to per capita population growth, which corresponds to the following model:
\begin{equation}
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,.
\end{equation}

@@

Using these Akaike weights, we calculated two important statistics.
First, the total Akaike weight of models with a lower {\ic} than the entirely ungrouped model; this is the probability that some model is more parsimonious than the fully ungrouped model.
Second, the total Akaike weight of models that used the same grouping for the rows as for the columns; this is the probability that the most parsimonious model grouped species equivalently in their competitive effects as in their competitive responses.
To evaluate the significance of the latter, we performed a Monte Carlo permutation test, comparing the total Akaike weight of those models that grouped rows and columns equivalently to the total Akaike weight of a equally sized random subset of the models.
%This is a test of whether the most parsimonious model is more likely to group the species equivalently in the rows as in the columns than it would be by chance.

@@

Overall, no pair of species can confidently be said to be coclassified together in the row or column groupings; the highest probability of classification was \SI[round-mode=figures,round-precision=3]{\Sexpr{max(weighted_col_coclassification_matrix - diag(num_species), weighted_row_coclassification_matrix - diag(num_species))*100}}{\percent} (\cref{fig:coclassification}).
Some pairs of species were far more likely to be coclassified than others, however, suggesting some structure to the best groupings.

>>>>

Dynamic models of ecological communities require many parameters, making them expensive to fit to experimental or observational data.
To reduce the number of parameters, species are often divided into groups \emph{a priori}, using phylogenetic or functional groups, and species within a group are assumed to behave identically.
Here, we assess the validity of such grouping by deriving the optimal groupings \emph{a posteriori}, from the parsimony of a Lotka--Volterra model fit to experimental data from six species of \emph{Drosophila} competing in laboratory conditions.
We find evidence that grouping can improve the parsimony of species interaction models, but we also observe that there is no single optimal grouping.
Further, those groupings that did prove beneficial could not be easily predicted from available species attributes \emph{a priori}, calling into question the prevailing method of \emph{a priori} species grouping.
Lastly, we find that species did not group equivalently in their competitive effect and in their response to competition, invalidating a key assumption often made implicitly when defining species groupings.
This suggests the need for a different, more data-ifnormed, approach to grouping in species interaction models.

@@

The Lotka--Volterra competition model in Eq.~(\ref{LVmodel}) is not linear in the parameters.
With the rows of the interaction matrix ungrouped, the regression could be reparameterised to be linear in the parameters by treating $r_i \alpha_{ij}$ as a single regression parameter for each $i,j$ pair.
With the rows grouped, however, this becomes impossible since each species still has an independent growth rate.
As such, we instead fit the models with the Gauss--Newton method for non-linear least squares and assumed a normal error distribution of the response variable.
To prevent heteroskedasticity and improve normality, we transformed total population growth to per capita population growth, which corresponds to the following model:
\begin{equation}
	\frac{N_{i,t+1} - N_{i,t}}{N_{i,t}} = r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,.
\end{equation}

@@

Using these Akaike weights, we calculated two important statistics.
First, the total Akaike weight of models with a lower {\ic} than the entirely ungrouped model; this is the probability that some model is more parsimonious than the fully ungrouped model.
Second, the total Akaike weight of models that used the same grouping for the rows as for the columns; this is the probability that the most parsimonious model grouped species equivalently in their competitive effects as in their competitive responses.
To evaluate the significance of the latter, we performed a Monte Carlo permutation test, comparing the total Akaike weight of those models that grouped rows and columns equivalently to the total Akaike weight of a equally sized random subset of the models.

@@

Overall, no pair of species can confidently be said to be coclassified together in the row or column groupings; the highest probability of classification was \SI[round-mode=figures,round-precision=3]{\Sexpr{max(weighted_col_coclassification_matrix - diag(num_species), weighted_row_coclassification_matrix - diag(num_species))*100}}{\percent} (\cref{fig:coclassification}).
However, some pairs of species were far more likely to be coclassified than others, suggesting some structure to the best groupings.
